### 2008-11-05 teoria_algoritmos/examen2/examen2.tex
limpieza del .tex de la tarea 2 de teoria de algoritmos

\documentclass{article}

\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage[spanish]{babel}

% Para codigo bonito
\newcommand{\singlespace}{\renewcommand{\baselinestretch}{1.0}}

\newcommand{\algoritmo}[2]
{
{
  \singlespace
  \begin{algorithm}[h]
    \caption{#1}
    \begin{algorithmic}[1]
      #2
    \end{algorithmic}
  \end{algorithm}
}
}

\newcommand{\theHalgorithm}{\arabic{algorithm}}

\lstset{mathescape=true}

\begin{document}

% Título

\title{Tarea 2 de Teoría de Algoritmos}
\author{Daniel Izquierdo \\ \#08-86809}
\date{6 de noviembre de 2008}

\maketitle

% Pregunta 1
\section{}

test

\end{document}