### 2009-05-28 bdhetero/tarea2/tarea2.tex
esqueleto de tarea 2 de bdhetero

\documentclass[12pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage[spanish]{babel}

\newcommand{\asig}{\ensuremath{\leftarrow}}

\begin{document}

% Título

\title{Bases de Datos Heterogéneas \\ Tarea 2}
\author{Daniel Izquierdo \\ \#08-86809}

\maketitle

\renewcommand{\labelenumi}{\alph{enumi}.}
\begin{enumerate}

% Pregunta 1
 \item

\end{enumerate}

\end{document}