### 2009-06-18 matematicas/tarea5.tex
prepare .tex de tarea 5 de matematicas

\documentclass{article}

\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage[spanish]{babel}
\usepackage{qtree}
\usepackage{enumerate}

% Para codigo bonito
\newcommand{\singlespace}{\renewcommand{\baselinestretch}{1.0}}
\newcommand{\piso}[1]{\left \lfloor #1 \right \rfloor}
\newcommand{\techo}[1]{\left \lceil #1 \right \rceil}
\newcommand{\stirlingU}[2]{\genfrac{[}{]}{0pt}{}{#1}{#2}}
\newcommand{\stirlingD}[2]{\genfrac{\{}{\}}{0pt}{}{#1}{#2}}

\newcommand{\algoritmo}[2]
{
{
  \singlespace
  \begin{algorithm}[h]
    \caption{#1}
    \begin{algorithmic}[1]
      #2
    \end{algorithmic}
  \end{algorithm}
}
}

\newcommand{\theHalgorithm}{\arabic{algorithm}}

\lstset{mathescape=true,frame=single}

\newcommand{\asig}{\ensuremath{\leftarrow}}
\renewcommand{\lstlistingname}{Listado} %listings en español

\begin{document}

% Título

\title{Tarea 5 de Matemáticas para Computación}
\author{Daniel Izquierdo \\ \#08-86809}
\date{25 de junio de 2009}

\maketitle

% Pregunta 1
\section{}

\end{document}