### 2009-07-11 bdhetero/tarea3/tarea3.tex
esqueleto correcto de tarea3 de bdhetero

\documentclass[10pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage[spanish]{babel}
\usepackage{qtree}

\newcommand{\asig}{\ensuremath{\leftarrow}}

\begin{document}

% Título

\title{Bases de Datos Heterogéneas \\ Tarea 3}
\author{Daniel Izquierdo \\ \#08-86809}

\maketitle

% Pregunta 1
\section{}

% Pregunta 2
\section{}

\end{document}