### 2009-07-11 bdhetero/tarea3/tarea3.tex
arbol de pregunta2 de tarea3 de bdhetero

\documentclass[10pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage[spanish]{babel}
\usepackage{qtree}

\newcommand{\asig}{\ensuremath{\leftarrow}}

\begin{document}

% Título

\title{Bases de Datos Heterogéneas \\ Tarea 3}
\author{Daniel Izquierdo \\ \#08-86809}

\maketitle

% Pregunta 1
\section{}

\begin{itemize}
 \item $v_1(x,y),v_6(x),v_7(y)$
 \item $v_1(y,x),v_4(x,y),v_6(x),v_7(y)$
 \item $v_3(y,x),v_3(x,y),v_6(x),v_7(y)$
\end{itemize}

% Pregunta 2
\section{}

% Pregunta 2.1
\subsection{}

\Tree
[.$\sigma_{\text{Premio.Premio}=\text{'Oscar'}, \text{Premio.Año}=2009}$
 [.$\Join_{Pelicula.CodPelicula=Premio.CodPelicula}$
   [.$\sigma_{\text{Evaluacion.Evaluacion}=\text{'Buena'}}$
    [.$\Join_{\text{Pelicula.CodPelicula}=\text{Evaluacion.CodPelicula}}$
     [.$\sigma_{\text{Pelicula.Director}=\text{'Almodovar'}}$
      Pelicula
     ]
     Evaluacion
    ]
   ]
   Premio
 ]
]

% Pregunta 2.2
\subsection{}

\end{document}