### 2009-10-10 ws2/reporte.tex
directorio de ws2, reporte de clase

\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage[spanish]{babel}

\newcommand{\tablaTripletas}[2]
{
\begin{tabular}{l|c}
 & Número de tripletas \\
 \hline
 Antes & #1 \\
 Después & #2 \\
\end{tabular}
}

\newcommand{\tablaPredicados}[1]
{
\begin{tabular}{l|c|c}
Predicado & Antes & Después \\
\hline
#1
\end{tabular}
}

\newcommand{\tablaInstancias}[1]
{
\begin{tabular}{l|c|c|c}
Clase & Antes & Después \\
\hline
#1
\end{tabular}
}

\newcommand{\resultadoDataSet}[5]
{
  \newpage

  \section{#1}

  \vspace{0.5cm}

  \tablaTripletas{#2}{#3}

  \vspace{1cm}

  \tablaPredicados{#4}

  \vspace{1cm}

  \tablaInstancias{#5}
}

\begin{document}

\resultadoDataSet{rdfPrimero}{89}{879}
{
    cursa & 14 & 70 \\
    dicta & 16 & 40 \\
    esProfesor & 0 & 27 \\
    esSupervisor & 9 & 45 \\
    esTutor & 0 & 36 \\
    rdf:type & 40 & 507 \\
    rdfs:subClassOf & 8 & 111 \\
    rdfs:subPropertyOf & 2 & 43 \\
}
{
    estudiante & 0 & 32 \\
    estudiantePostGrado & 4 & 20 \\
    estudiantePostGradoOcasional & 4 & 20 \\
    estudiantePreGrado & 4 & 20 \\
    materia & 0 & 32 \\
    materiaPostGrado & 4 & 20 \\
    materiaPreGrado & 4 & 20 \\
    persona & 0 & 40 \\
    profesor & 4 & 20 \\
    rdfs:Class & 11 & 66 \\
    rdfs:Property & 5 & 37 \\
    rdfs:Resource & 0 & 180 \\
}

\resultadoDataSet{rdfSegundo}{120}{1193}
{
    cursa & 14 & 70 \\
    dicta & 16 & 40 \\
    esProfesor & 0 & 54 \\
    esSupervisor & 18 & 90 \\
    esTutor & 0 & 72 \\
    rdf:type & 62 & 713 \\
    rdfs:subClassOf & 8 & 111 \\
    rdfs:subPropertyOf & 2 & 43 \\
}
{
    estudiante & 0 & 32 \\
    estudiantePostGrado & 4 & 20 \\
    estudiantePostGradoOcasional & 24 & 120 \\
    estudiantePreGrado & 4 & 20 \\
    materia & 0 & 40 \\
    materiaPostGrado & 5 & 25 \\
    materiaPreGrado & 5 & 25 \\
    persona & 0 & 40 \\
    profesor & 4 & 20 \\
    rdfs:Class & 11 & 66 \\
    rdfs:Property & 5 & 37 \\
    rdfs:Resource & 0 & 268 \\
}

\resultadoDataSet{rdfTercero}{162}{1489}
{
    cursa & 14 & 70 \\
    dicta & 16 & 40 \\
    esProfesor & 0 & 108 \\
    esSupervisor & 36 & 180 \\
    esTutor & 0 & 144 \\
    rdf:type & 86 & 793 \\
    rdfs:subClassOf & 8 & 111 \\
    rdfs:subPropertyOf & 2 & 43 \\
}
{
    estudiante & 0 & 32 \\
    estudiantePostGrado & 4 & 20 \\
    estudiantePostGradoOcasional & 48 & 160 \\
    estudiantePreGrado & 4 & 20 \\
    materia & 0 & 40 \\
    materiaPostGrado & 5 & 25 \\
    materiaPreGrado & 5 & 25 \\
    persona & 0 & 40 \\
    profesor & 4 & 20 \\
    rdfs:Class & 11 & 66 \\
    rdfs:Property & 5 & 37 \\
    rdfs:Resource & 0 & 308 \\
}

\end{document}