### 2010-01-12 tesis/propuesta/traduccion/traduccion.tex

\documentclass{article}

\usepackage[utf8x]{inputenc}
\usepackage[spanish]{babel}

\begin{document}

\tableofcontents

\end{document}